---
title: "Gapminder-dplyr"
output: html_document
---

# Loading required libraries
```{r load packages}
library(tidyverse)
```

# Load data
```{r load data}
gapminder <- read.csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")
```

# Look at the data
```{r look at data}
head(gapminder)
tail(gapminder)
str(gapminder)
names(gapminder)
ncol(gapminder)
```

# Look at individual variables within the df
```{r}
gapminder$lifeExp
gapminder$country
```

# Working with dplyr (section 6.4 in the workbook)

## Filter() function. This operates on rows
```{r}
filter(gapminder, lifeExp < 40) # Filter by age
filter(gapminder, country == "Mexico") # Filter by country
filter(gapminder, country == "Mexico", year == 2002) # Filter by country and year
brazil_life_exp <- filter(gapminder, country == "Brazil", year > 1986 & year < 2008)
mean(brazil_life_exp$lifeExp)
```

## Select() function. This operates on columns
```{r}
select(gapminder, year, country, lifeExp) # Note it produces the order that you ask for. The first item must be the dataset though.
select(gapminder, -continent, -gdpPercap) # The first again is the df. The others are ones that you don't want
```

## Using filter() and select() together (Note shortcut fo arrow is Option minus)
```{r}
# Create a cambodia specific df without continent and lifeExp
gap_cambodia <- filter(gapminder, country == "Cambodia") #df1
gap_cambodia2 <- select(gap_cambodia, -continent, -lifeExp) #df2
```

## Using pipe operator
Note - command shift m gives the pipe ('and then') on a mac %>% 
note - using pipe is a nice way to avoid nested functions

```{r}
gapminder %>% head()
```
The pipe takes whatever it to its left and feeding it to whatever is on the right
So the above takes gapminder and feeds it to the head function

```{r}
gap_cambodia <- filter(gapminder, country == "Cambodia") %>% #and then
  select(-continent, -lifeExp)
```

Another option
```{r}
gap_cambodia <- gapminder %>% #and then
  filter(country == "Cambodia") %>% #and then
  select(-continent, -lifeExp)
```


## Using mutate()
```{r}
gapminder %>%  
  mutate(gdp = pop * gdpPercap) 
# You don't need to press enter after the pipe but it is a convention to make it more readable
#gdp is a new column. Note no == as it isn't a question but a command
 
```

Calculate the population in thousands for all Asian countries in the year 2007
```{r}
gapminder %>% 
  filter(continent == "Asia",year == 2007) %>% 
  mutate(popthousands = pop/1000)
```

## Using group_by()
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  mutate(cont_pop = sum(pop)) # This is not great. See next option as better option
```

Add in summarize()
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  summarize(cont_pop = sum(pop))
```

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop))
```

Use ungroup (to undo the grouping as once the grouping is applied it will apply the whole way)
```{r}
df1 <- gapminder %>% 
  group_by(continent, year) %>% 
  mutate(cont_pop = sum(pop)) %>%
  ungroup() # Just good practice to ungroup just in case it stays grouped & causes problems
```

## Using arrange().
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(cont_pop = sum(pop)) %>% 
  arrange(-year) # Arrange by year in descending order
```

```{r}
gapminder %>% 
  mutate(gdp = pop * gdpPercap) %>% 
  group_by(continent) %>% 
  mutate(maxGDP = max(gdp)) %>% 
  filter(gdp == maxGDP)
  ungroup()
```




