
# Learning ggplot2
We're learning ggplot2. It's going to be amazing

Load tidyverse
```{r tidyverse}
library(tidyverse)
```

Note that the "R data" title in the R chunk runs R but the data part is actually a sub-heading. If you then click on the learning ggplot2 at the bottom on this window you can see all of the headings and the chunk titles

Get the data from github and get a summary

```{r data}
ohid_data <- read.csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")
summary(ohid_data)
describe(ohid_data)
dim(ohid_data)
```

We start with ggplot. Note data need to be in a dataframe
The first ggplot section outlines which variables to plot
The geom section says which type of plot is needed
geom_point is a scatter plot
geom_jitter is a scatter plot with jittering
geom_bar is a barplot
geom_histogram is a histogram

```{r}
ggplot(data = ohid_data, aes(x = HDI)) + 
  geom_histogram()
```
Multiple datasets can be overlaid onto each other
Get second dataset

Note the ggplot is the global assignment - it will be used unless a later geom assigns a different object

Note that aes means aesthetics
```{r multiple geoms}
OHI_summary <- read.csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_scores_georegion_summary.csv")

ggplot(data = OHI_summary, aes(x = georegions, y = OHI_score_average)) +
  geom_bar(stat = "identity") +
  geom_jitter(data = ohid_data, aes(x = georegion_one, y = OHI_score))
```

These two actually produce the same graph
```{r Different ways to specify the ggplot}
ggplot(data = ohid_data, aes(y=OHI_score, x = HDI, color=georegion_one)) +
  geom_point()
  
ggplot(data = ohid_data) +
  geom_point(aes(y = OHI_score, x = HDI, color=georegion_one)) 
```

# Aes notes
color
fill
label
linetype
shape
alpha
size

# Adding a third variable with aes
This plot makes the dots bigger or smaller in line with coastal population size
```{r}
ggplot(data = ohid_data, aes(x = OHI_score, y = HDI, size = coastal_pop)) +
  geom_point()
```

This plot makes the dots coloured in line with coastal population size (a continuous variable)
```{r}
ggplot(data = ohid_data, aes(x = OHI_score, y = HDI, color = coastal_pop)) +
  geom_point()
```

This plot makes the dots coloured in line with a discrete variable
```{r}
ggplot(data = ohid_data, aes(x = OHI_score, y = HDI, color = georegion_one)) +
  geom_point()
```

This plot changes dot shape in line with a discrete variable
```{r}
ggplot(data = ohid_data, aes(x = OHI_score, y = HDI, shape = georegion_one)) +
  geom_point()
```

# Themes
Themes provide default plotting settings
```{r}
ggplot(data = ohid_data, aes(x = OHI_score, y = HDI, shape = georegion_one)) +
  geom_point() +
  theme_bw()
```

```{r}
source('https://raw.githubusercontent.com/OHI-Science/ohiprep/master/src/R/scatterTheme.txt')       

ggplot(data = ohid_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() + 
  scatterTheme
```

Change labels
```{r}
ggplot(data = ohid_data, aes(y = OHI_score, x = HDI, color = georegion_one)) +
  geom_point() +
  labs(y = "OHI score, 2017",
       x = "Human Development Index",
       title = "Human development is good for oceans",
       color = "Georegions")
```

To use RColourBrewer
```{r}
library(RColorBrewer)
display.brewer.all()

my_palatte <- brewer.pal(n = 9, "YlOrRd")
my_palatte
```
Look at "Hint 3" in the manual for some colour hints depending on whether the variable is continuous or discrete

```{r}
ggplot(data = ohid_data, aes(y = OHI_trend, x = OHI_score, color = HDI)) +
  geom_point(size = 3) +
  scale_color_gradientn(colors = my_palatte)
```

Can specify the colours in a gradient if variable is continuous
```{r}
ggplot(data = ohid_data, aes(y = OHI_trend, x = OHI_score, color = HDI)) +
  geom_point(size = 3) +
  scale_color_gradientn(colors = c("blue","red"))
```

If variable is discrete
```{r}
ggplot(data = ohid_data, aes(y = HDI, x = OHI_score, color = georegion_one)) +
  geom_point(size = 3) +
  scale_color_manual(values = my_palatte)
```

